### 2022-03-31 outputs/paper/paper.Rmd
added graphs

---
title: "Looking at the India State Findings from the 1992 DHS Final Report"
subtitle: ""
author: 
  - Bilal Haq
  - Ritvik Puri
thanks: "Code and data are available at: https://github.com/haqbilal/India-1992-DHS-State-Findings"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

dhs <- readr::read_csv(here::here("inputs/data/cleaned_data.csv"))
dhs <- dhs %>%
  select(states, everything())
view(dhs)
```

# Introduction

# Data
Graphs
```{r}
# Illiterate females get married early
dhs %>%
  ggplot(aes(`Percent of illiterate females`, `Percent of women married before age 18`, color = states)) +
  geom_point()

# Illiteracy has no connection to contraceptives?
dhs %>%
  ggplot(aes(`Percent of illiterate females`, `Percent of women using any contraceptive method`)) +
  geom_point() +
  geom_smooth(method = "lm")

# Women going to school in each state
dhs %>%
  ggplot(aes(x=states, y=`Percent of females attending schools`)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_bw()
  #theme(axis.text.x = element_text(angle = 90))

#Unmet family planning leading to infant mortality are apparently not related
dhs %>%
  ggplot(aes(`Unmet need for family planning`, `Infant mortality rate`)) +
  geom_point()
```

# Results

# Discussion

include details of the methodology used by the DHS you used, and its key features, strengths, and weaknesses.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References