### 2022-04-01 outputs/paper/paper.Rmd
filtered out unneeded vars

---
title: "Now that's fresh water: Trends in India from the 1992 Demographic and Health Survey"
subtitle: ""
author: 
  - Bilal Haq
  - Ritvik Puri
thanks: "Code and data are available at: https://github.com/haqbilal/India-1992-DHS-State-Findings"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Data

# Results

# Discussion

include details of the methodology used by the DHS you used, and its key features, strengths, and weaknesses.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References